### 2010-10-27 science_ms.tex
Added some details about the simulation.  Revised after another proofreading.  Removed a paragraph in the discussion.  Within 80 words of the word limit, including the acknowledgements.  I think this should be fine.

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

\usepackage{graphicx}

\newcommand\aap{{A\&A}}% 
         % Astronomy and Astrophysics 
\newcommand\aaps{{A\&AS}}% 
         % Astronomy and Astrophysics, Supplement 
\newcommand\aj{{AJ}}% 
         % Astronomical Journal 
\newcommand\araa{{ARA\&A}}% 
         % Annual Review of Astron and Astrophys 
\newcommand\apj{{ApJ}}% 
         % Astrophysical Journal 
\newcommand\apjl{{ApJ}}% 
         % Astrophysical Journal, Letters 
\newcommand\apjs{{ApJS}}% 
         % Astrophysical Journal, Supplement 
\newcommand\mnras{{MNRAS}}% 
         % Monthly Notices of the RAS 
\newcommand\nat{{Nature}}% 
         % Nature 
\newcommand\pasp{{PASP}}% 
         % Publications of the ASP 
\newcommand{\cubecm}{\ifmmode{{\rm cm^{-3}}}\else{cm$^{-3}$}\fi}
\newcommand{\lsim}{\lower0.3em\hbox{$\,\buildrel <\over\sim\,$}}
\newcommand{\gsim}{\lower0.3em\hbox{$\,\buildrel >\over\sim\,$}}
\newcommand{\kms}{\ifmmode{{\rm km~s^{-1}}}\else{km s$^{-1}$}\fi}
\newcommand{\hh}{H$_2$}
\newcommand{\Ms}{\ifmmode{M_\odot}\else{$M_\odot$}\fi}
\newcommand\tento[1]{$10^{#1}$}
\newcommand{\tvir}{\ifmmode{T_{\rm{vir}}}\else{$T_{\rm{vir}}$}\fi}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

%% We will uncomment this later
\usepackage{scicite}
\bibliographystyle{Science}

% Use times if you have the font installed; otherwise, comment out the
% following line.

\usepackage{times}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

\title{The Birth of a Galaxy: Population III Metal Enrichment and
  Population II Stellar Populations}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{John H.~Wise,$^{1,2\ast}$ Matthew J.~Turk$^{3}$, Michael
  L.~Norman,$^{3}$ Tom Abel$^{4}$\\
\\
\normalsize{$^{1}$Department of Astrophysical Sciences, Princeton
University,}\\
\normalsize{Peyton Hall, Ivy Lane, Princeton, NJ 08544}\\
\normalsize{$^{2}$Hubble Fellow}\\
\normalsize{$^{3}$Center for Astrophysics and Space Sciences,}\\
\normalsize{University of California at San Diego, La Jolla, CA 92093}\\
\normalsize{$^{4}$Kavli Institute for Particle Astrophysics and
  Cosmology, }\\
\normalsize{Stanford University, Menlo Park, CA 94025}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:
jwise@astro.princeton.edu}
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
  Supernova explosions from the first stars in the universe, which are
  predicted to be very massive, chemically enriched the intergalactic
  medium, originally a primordial mix of hydrogen and helium.  A
  fraction of these metals propagate into star-forming regions,
  providing more efficient cooling mechanisms and, in turn, increasing
  gaseous fragmentation in the first stellar clusters.  Here, we
  report on the birth of two high-redshift dwarf galaxies in a
  cosmological radiation hydrodynamics simulation.  We find that only
  one supernova is sufficient to enrich the nearby star-forming gas
  with a floor of 1/1,000 of the solar metallicity ratio, preventing
  any further metal-free star formation within 10 kpc of the past
  explosion.  This provides a natural explanation for the observed
  floor on damped Lyman alpha (DLA) systems metallicities reported in
  the literature, which is of this order.  We follow the entire star
  formation history of these galaxies, providing robust models to
  interpret the stellar populations of nearby dwarf galaxies and
  future observations of high-redshift dwarf galaxies.
\end{sciabstract}

% In setting up this template for *Science* papers, we've used both
% the \section* command and the \paragraph* command for topical
% divisions.  Which you use will of course depend on the type of paper
% you're writing.  Review Articles tend to have displayed headings, for
% which \section* is more appropriate; Research Articles, when they have
% formal topical divisions at all, tend to signal them with bold text
% that runs into the paragraph, for which \paragraph* is the right
% choice.  Either way, use the asterisk (*) modifier, as shown, to
% suppress numbering.

\section*{Motivation}

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \cite{ABN02, Bromm02_P3, Yoshida03, OShea07a}.  A fraction of
these stars enrich the surrounding intergalactic medium (IGM) when
they go supernova, which can happen in stars $\lsim 40~\Ms$ in Type II
supernovae (SNe) or in stars roughly between 140 \Ms~and 260 \Ms~in
pair-instability SNe (PISNe; \cite{2002ApJ...567..532H}).  The host
halo and the neighboring halos are then enriched with this ejecta.
There exists a critical metallicity that is $\sim 10^{-6} Z_\odot$ if
dust cooling is efficient \cite{Omukai05, Schneider06_Frag, clark08}
and $\sim 10^{-3.5} Z_\odot$ otherwise \cite{Bromm01,
  2009ApJ...691..441S}, where the gas can cool rapidly, lowering its
Jeans mass.  An intermediate characteristic mass of $\sim10~\Ms$ can
be occur if the gas cooling is suppressed to the cosmic microwave
background (CMB) temperature \cite{Larson98, Tumlinson07_IMF,
  2009ApJ...691..441S}.  The resulting Population II star cluster will
thus have a lower characteristic stellar mass than its metal-free
progenitors.  These first stellar clusters may be connected to stars
in the Milky Way halo and nearby dwarf spheroidal (dSph) galaxies,
both with a metallicity floor of [Z/H]\footnote{We use the
  conventional notation, [Z/H] $\equiv$ log(Z/H) -
  log(Z$_\odot$/H$_\odot$)} = --4 \cite{Beers05, Tafelmeyer10,
  Frebel10_Obs}.

The transition from Pop III to Pop II star formation (SF) is solely
dependent on the propagation of metals from the SNe remnants into
future sites of SF.  Their flows are complex because of the
interactions between the SN blastwave, cosmological accretion and halo
mergers, and nearby stellar feedback.  In minihalos ($M \sim
10^6~\Ms$), radiation from a massive Pop III star can drive a 30
\kms~shock, which is 10 times greater than the escape velocity of the
halo, and leaves behind a warm ($3 \times 10^4$ K) and diffuse (0.5
\cubecm) medium \cite{Kitayama04, Whalen04, Abel07}.  This aids in the
expansion of the blastwave because it delays the transition to the
Sedov-Taylor and snowplow phases.  In PISNe, approximately half of the
metals stay in the IGM with a metal bubble size of a few kpc
\cite{Wise08_Gal, Greif10}.  The blastwave may induce SF in nearby
halos through the compression of the gas \cite{Ferrara98}, and
timescales for metal mixing into the dense gas are many dynamical
times \cite{Cen08} for shock velocities $\lsim100~\kms$.

Numerical simulations are useful to detangle and study these
complexities and the transition from Pop III to II stars.  In this
Report, we present the simulations that include both types of SF, and
their radiative and mechanical feedback.  The methods used here
incorporate and link together recent results from metal-enriched and
metal-free star formation, the critical metallicity, and
pair-instability supernovae.  This is the first time it has been
possible to link the formation and feedback of the first stars to
protogalaxies, resolving the important scales and including the most
important physical effects.

\section*{Results}
\label{sec:results}

Here we present the gaseous and stellar evolution of two selected
halos in the simulation: one that has an early mass buildup but no
major mergers after $z=12$, and one that experiences a series of major
mergers between $z=10$ and $z=7$.  We name the halos ``quiet'' and
``intense'', respectively.  The entire simulation contains 38 galaxies
with 3640 Pop II stellar clusters and captures the formation of 333
Pop III stars.  The halo mass function and galaxy luminosities are
plotted in Figure \ref{fig:massfn}.  This simulation has $1.4 \times
10^8$ computational elements and a maximal spatial resolution of 0.1
pc.

% We start with the mass accretion history and enrichment of the
% halos.  We then compare the nature of the stellar populations in
% these halos.

We illustrate the state of the simulation at $z=7$ in Figure
\ref{fig:projections} with density weighted projections of gas
density, temperature, and metallicity, showing the entire box and
focusing on the two halos of interest.  Radiative and mechanical
feedback create a multi-phase medium inside these halos, which are
embedded in a warm and ionized IGM.

\subsection*{Mass evolution}
\label{sec:halo}

%\li Describe the evolution of the baryon fraction and metallicity from
%PISN and Type II SNe metals in the two halos.

Figure \ref{fig:evo}a shows the total, metal-enriched stellar, and gas
mass history of the most massive progenitors of both halos.  The quiet
halo undergoes a series of major mergers at $z > 12$, growing by a
factor of 30 to $2.5 \times 10^7 \Ms$ within 150 Myr.  Afterwards it
only grows by a factor of 3 by $z=7$ mainly through smooth accretion
from the filaments and IGM.  It is the most massive halo in the
simulation between redshifts 13 and 10.  At the same time, the intense
halo has a mass $M = 3 \times 10^6 \Ms$, but it is contained in a
biased region on a comoving scale of 50 kpc with $\sim25$ halos with
$M \sim 10^6 \Ms$.  After $z=10$, these halos hierarchically merge to
form a $10^9 \Ms$ halo at $z=7$ with two major mergers at redshifts 10
and 7.9, seen in the rapid increases in total mass in Figure
\ref{fig:evo}a.  The merger history of the two halos are not atypical
as dark matter halos can experience both quiescent and vigorous mass
accretion rates.

Both halos start forming Pop II stars when $M = 10^7 \Ms$.  This is
consistent with the filtering mass $M_f$ of high-redshift halos when
it accretes mainly from a pre-heated IGM \cite{gnedin98, gnedin00,
  Wise08_Gal}.  Afterwards these halos can cool efficiently through
\hh~cooling, sustaining constant and sometimes bursting SF.  The
latter characteristics are equated with the definition of a galaxy.
The quiet halo forms $10^5 \Ms$ of stars by $z=9$.  This initial
starburst photo-evaporates the majority of its molecular clouds, in
addition to heating and ionizing the surrounding IGM out to a radius
of 10--15 kpc at $z=9$.  These respectively reduce the in-situ and
external cold gas supply that could feed future SF.

The gas fractions of both halos decrease from 0.15 to 0.08 by outflows
driven by ionization fronts and blastwaves in their initial
starbursts.  The quiet halo never recovers to the cosmic fraction
$\Omega_b/\Omega_M$ for the following reason.  The lack of major
mergers of halos above the filtering mass leads to a small gas
fraction because halos with $M < M_f$ are likely photo-evaporated,
hosting diffuse warm gas reservoirs instead of cold dense cores.
After $z=10$, the halo mainly accretes warm diffuse gas from the
filaments and IGM.  In contrast, the intense halo grows from major
mergers of halos with $M > M_f$.  The progenitor halos involved in the
major mergers are able to host molecular clouds and have higher gas
fractions.  Between $z=10$ and $z=8$, the gas fraction increases from
0.07 to 0.12 until it jumps to 0.14 when a gas-rich major merger
occurs.  The stellar mass accordingly increases with the ample supply
of gas during this period.  The gas fraction then decreases by 0.01
when galactic outflows are generated from the starburst associated
with this major merger.

\subsection*{Metallicity evolution}
\label{sec:zevo}

The evolution of the stellar and gas metallicity of both halos are
illustrated in Figure \ref{fig:evo}b.  PISNe from Pop III stars enrich
the nearby IGM out to a radius of 10 kpc and provides a metallicity
floor of $[Z_3/H] \sim -3$.  The metallicity from Pop II SNe initially
enrich the ISM of both halos to an average $[\bar{Z}_2/H]$ between
--3.5 and --3.

In the quiet halo, an equilibrium of $[\bar{Z}_2/H] \sim -2.5$ is
established between metal-rich outflows and metal-poor inflows.
Galactic outflows are directed in the polar directions of the gas
disk, keeping the adjacent filaments metal-poor.  These features and a
well-mixed ISM (cf.~\cite{Wise08_Gal, Greif10}) are apparent in the
metallicity projections in Figure \ref{fig:projections}.  The average
stellar metallicity is within 0.5 dex of $[\bar{Z}_2/H]$.

In the intense halo, the first few Pop II star clusters have $[Z/H]$
between --1 and --2 and dominate the average stellar metallicity at $z
> 8$.  Afterwards the metallicity increases by a factor of 30 to
$[\bar{Z}_2/H] = -1.5$ through self-enrichment from a starburst.
Because this halo is located in a large-scale overdensity, most of the
ejecta falls back into the halo after reaching distances up to 20
comoving kpc, keeping the halo gas metallicity high because the
inflows are relatively metal-rich themselves.  After the $z \sim 8$
starburst, the average stellar metallicity follows the average gas
metallicity within 0.1 dex.

\subsection*{Population III star formation history}

The most massive progenitor of the quiet halo interestingly never
hosts a Pop III star.  Instead a nearby halo forms a Pop III star,
which (randomly) produces a PISNe at $z=16$.  The blastwave overruns
the most massive progenitor, and the dense core survives this event
and is enriched by this PISN, triggering the transition to Pop II SF.
Other progenitors host three Pop III stars, forming at $z = 15.4,
14.2, 13.8$, with the latter producing a PISN.  Metal enrichment from
these two PISNe and Pop II SF quench Pop III SF in this halo.  The
progenitors of the intense halo host a total of 56 Pop III stars with
21 producing PISNe.  The first forms in a $6 \times 10^5 \Ms$ halo at
$z=19$.  Pop III stars form on a regular interval in the halo's
progenitors until $z=9$ when most of these halos enter the metal-rich
bubble surrounding the intense halo.

% There are two stars forming at $z=7.7$ with [Z/H] = --5.7, whose
% molecular clouds were enriched by external blastwaves that only
% partially enriched the dense regions.  This falls into the metallicity
% range where fragmentation could occur if dust cooling is efficient
% below \tento{-4} $Z_\odot$.  The most massive progenitors of the
% intense halo cease to form Pop III stars at $z<12$ because of nearby
% or intrinsic PISNe.

\subsection*{Population II stellar populations}
\label{sec:pop}

%% \li Present results on the star formation history and the physical
%% reasons for the features seen in the SFH and metallicities.
%%
%% \li Present the stellar metallicity distribution at the final time

Figure \ref{fig:pops} shows the SF history (SFH), metallicity
distribution, and SF rates of both halos.  A nearby PISNe provides a
metallicity floor of [Z/H] = --2.8 in the quiet halo at which
metallicity the first Pop II stars form.  The stellar metallicity
evolution exhibits what is expected from an isolated system with the
stellar feedback steadily enriching the ISM, resulting in a
correlation between stellar age and metallicity.  After $z=10$ the
metallicities plateau at [Z/H] = --2.1 for reasons previously
discussed.  The SFR peaks at $z=10$ and decreases as the cold gas
reservoir is depleted.  Around $z=7.5$, a 25:1 minor merger occurs,
and the gas inside the satellite halo is compressed, triggering
metal-poor, [Z/H] = --3.2, SF during its nearest approach.  This halo
remains metal-poor because most of metal enrichment in the quiet halo
occurs in bi-polar flows perpendicular to the galaxy disk and
filament.  Stars with [Z/H] $<$ --3 compose 1.6 percent of the total
stellar mass.

In contrast with the quiet halo, the intense halo undergoes a few
mergers of halos with an established stellar population.  This creates
a superposition of age-metallicity tracks in the SFH, seen in the
complexity of Figure \ref{fig:pops}.  The first two Pop II stellar
clusters have an unexpectedly high metallicity [Z/H] $\sim$ --1, which
occurs when a PISN blastwave triggers SF in two neighboring halos.
Most of the early SF have [Z/H] = --2.5.  At $z=9$, the halo's virial
temperature reaches \tento{4} K.  This combined with a 10:1 merger
creates a starburst that quickly enriches the halo to [Z/H] = --1.5 by
$z=8$.  The halo continues to enrich itself afterwards.  The spikes in
the scatter plot correspond to SN triggered SF in nearby molecular
clouds that are enriched up to a factor of 10 with respect to the ISM.
However their mass fraction are small compared to the total stellar
mass.  The starburst at $z=9$ creates a bimodal metallicity
distribution with peaks at [Z/H] = --2.4 and --1.2 with the metal-rich
component mainly being created after the starburst.  Two systems with
sizable stellar components merge into the halo at $z \sim 8$, and
their stellar populations are still discernible in the metallicity-age
plot.  Stars with [Z/H] $<$ --3 compose 1.8 percent of the total
stellar mass.

\section*{Discussion and Conclusions}

In this Report, we focus on the birth of two galaxies prior to
reionization with a cosmological AMR radiation hydrodynamics
simulation.  Supernovae from Pop III stars provide the necessary heavy
elements for the transition to a Population II stellar population,
which we have directly simulated.  These two galaxies have a 10--15\%
probability in surviving as present-day ``fossil'' galaxies
\cite{Gnedin06}, otherwise they will be incorporated into galactic
stellar halos.  A $z=8.55$ galaxy was recently spectroscopically
confirmed that is contained in an ionized bubble with radius 0.1--0.5
Mpc with an uncharacterized population of galaxies contributing to the
local ionizing radiation field \cite{Lehnert10_z8.6}.  The galaxies
simulated here might represent this population.

We find that one PISN is sufficient to enrich the star-forming halo
and surrounding $\sim 5$ kpc to a metallicity of 10$^{-3} Z_\odot$,
given $M_{\rm char} = 100~\Ms$.  DLA systems have a metallicity floor
on the same order \cite{Wolfe05_Review}, and metal enrichment from Pop
III SNe provides a possible explanation.

If the first stars have a lower characteristic mass that favor
hypernovae \cite{Tumlinson07_IMF}, then this metallicity floor should
be lowered by a factor of $\sim 10$ because (1) the metal ejecta is
lowered by a factor of $\sim 50$ and (2) the mixing mass is
approximately decreased by a factor of $(E_{\rm hyp}/E_{\rm
  PISN})^{3/5}$ in the Sedov-Taylor solution, where $(E_{\rm
  hyp}/E_{\rm PISN}) \sim 0.1$ is the ratio of explosion energies of a
hypernova and PISN.  In the case where this metallicity floor is less
than the critical metallicity, then the next instance of SF will
further enrich the ISM, solidifying the transition to Pop II SF.  We
conclude that it only takes one, at most two, SNe from Pop III stars
in the halo progenitors to complete the transition to Population II
\cite{Frebel10}.  The question of whether the critical metallicity is
\tento{-6} or \tento{-3.5} $Z_\odot$ is most applicable to nearby
halos where the heavy elements mix slowly into dense cores as the
blastwave overtakes it.  Less than two percent of stars have [Z/H] $<
-3$ in both systems, consistent with observations of metal-poor stars
in the halo and dSphs \cite{Beers05, Battaglia10}.

% We also find that the merger history plays an important role in
% supplying gas into halos after its first epoch of SF.  Mergers of
% halos below the filtering mass are inefficient in providing gas
% whereas the opposite is true for merging halos above the filtering
% mass.  Halos do not necessarily need $\tvir \ge 10^4$ K to form
% significant stellar populations; however SFRs dramatically increase,
% and thus metal enrichment, when this threshold is reached.  Halos
% with mostly gas-poor mergers or a quiet merger history result in a
% monotonic increase in metallicity with time.  SFHs become more
% complex with multiple metallicity-age tracks if the halo experiences
% mergers with other halos that have an established stellar
% population.  Furthermore SF that is triggered by blastwaves
% interacting with molecular clouds can have metallicities up to a
% factor of 10 higher than the main starburst.  Starbursts at $\tvir =
% 10^4$ K can enrich the host halo enough to create a bimodal
% metallicity distribution, where the metal-poor component is created
% before the burst.  Note that mergers of stellar populations can also
% create a similar bimodal distribution, which are observed in dSphs
% \cite{Battaglia10}.

We have shown that it is possible to simulate the formation of a
high-redshift dwarf galaxy and its entire SFH with radiative and
mechanical feedback.  These results provide invaluable insight on the
first galaxies and the role of metal-free stars in the early universe.
There exists a wealth of information in this simulation, and we plan
to follow up this preliminary report with more detailed analysis of
the metal enrichment of the IGM, global SF rates, and observational
connections with high-redshift galaxies and Local Group dwarf galaxies
in the near future.

% Your references go at the end of the main text, and before the
% figures.  For this document we've used BibTeX, the .bib file
% scibib.bib, and the .bst file Science.bst.  The package scicite.sty
% was included to format the reference numbers according to *Science*
% style.

\bibliography{ms}

% Following is a new environment, {scilastnote}, that's defined in the
% preamble and that allows authors to add a reference at the end of the
% list that's not signaled in the text; such references are used in
% *Science* for acknowledgments of funding, help, etc.

\begin{scilastnote}
\item Support for this work was provided by NASA through Hubble Fellowship
grant \#120-6370 awarded by the Space Telescope Science Institute,
which is operated by the Association of Universities for Research in
Astronomy, Inc., for NASA, under contract NAS 5-26555.  Computational
resources were provided by NASA/NCCS award SMD-09-1439.
J.~H.~W. thanks Renyue Cen, Amina Helmi, Marco Spaans, and Eline
Tolstoy for enlightening discussions.
\end{scilastnote}

% For your review copy (i.e., the file you initially send in for
% evaluation), you can use the {figure} environment and the
% \includegraphics command to stream your figures into the text, placing
% all figures at the end.  For the final, revised manuscript for
% acceptance and production, however, PostScript or other graphics
% should not be streamed into your compliled file.  Instead, set
% captions as simple paragraphs (with a \noindent tag), setting them
% off from the rest of the text with a \clearpage as shown  below, and
% submit figures as separate files according to the Art Department's
% instructions.

\clearpage
\begin{figure*}
\includegraphics[width=0.6\textwidth]{Lstar.eps}
\caption{\label{fig:massfn} \textbf{Top}: Halo mass function of the
  simulation compared with the analytical fit (solid) of
  \cite{Warren06}.  \textbf{Bottom}: Total Pop II luminosities of
  halos with constant mass-to-light ratios overplotted.}
\end{figure*}

\clearpage
\begin{figure*}
\includegraphics[width=\textwidth]{f1.eps}
\caption{\label{fig:projections} Density-weighted projections of gas
  density (top), temperature (middle), and metallicity (bottom) at
  $z=7$.  The left column shows the entire simulation volume, where
  the center and right columns focus on the intense and quiet halos,
  which are marked by left and right arrows in the upper-left panel.
  The metallicity projections are a composite picture of metals
  originating from Pop III (red) and Pop II (blue) stars with magenta
  indicating a mixture of the two.}
\end{figure*}

\clearpage
\begin{figure*}
% Need to include f2.eps
\includegraphics[width=\textwidth]{f2.eps}
\caption{\label{fig:evo} (a) Evolution of the total halo mass (top),
  stellar mass (middle), and gas fraction (bottom) of the quiet
  (dashed) and intense (solid) halos.  (b) Mass-weighted stellar
  metallicities and gas metallicities enriched by Pop II and Pop III
  SNe of the intense (top) and quiet (bottom) halos.}
\end{figure*}

\clearpage
\begin{figure*}
\includegraphics[width=\textwidth]{f3.eps}
\caption{\label{fig:pops} The scatter plots show the SF history of the
  quiet (left) and intense (right) halos as a function of metallicity
  at $z=7$.  Each circle represents a star cluster, whose area is
  proportional to its mass.  The open circles in the upper right
  represent sizes of $10^3$, $10^4$, and $10^5$ \Ms~star clusters.
  The dashed lines in the right panel guide the eye to two stellar
  populations that were formed in two satellite halos, merging at
  $z=7.5$.  The upper histogram shows the SF rate.  The right
  histogram depicts the stellar metallicity distribution.}
\end{figure*}

%\noindent {\bf Fig. 1.} Please do not use figure environments to set
%up your figures in the final (post-peer-review) draft, do not include graphics in your
%source code, and do not cite figures in the text using \LaTeX\
%\verb+\ref+ commands.  Instead, simply refer to the figure numbers in
%the text per {\it Science\/} style, and include the list of captions at
%the end of the document, coded as ordinary paragraphs as shown in the
%\texttt{scifile.tex} template file.  Your actual figure files should
%be submitted separately.

%\include{som}

\end{document}