### 2011-01-22 table1.tex
Finished a draft of halo properties evo. section.  Added table with halo z=7 properties.

\begin{table*}
\begin{minipage}{13cm}
\caption{Halo properties of the 10 most massive halos at $z=7$ in the
  fidicual model}
\label{tab:halos}
\begin{tabular}{@{}cccccccccccc}
\hline

Halo & $M_{\rm halo}/\Ms$ & $a_c$ & $f_{\rm gas}$ & $M_\star/\Ms$ & $N_{\rm
  III}$ & $M/L$ & [Z$_2$/H] & [Z$_3$/H] & $\lambda_{\rm DM}$ &
$\lambda_{\rm g}$ \\

(1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) \\

\hline

0 & $1.0 \times 10^9$ & & 0.135 & $2.3 \times 10^6$ & 21 & 6.5 & -1.3 &
-2.7 & 0.046 & 0.055 \\

1 & $1.5 \times 10^8$ & & 0.111 & $1.5 \times 10^5$ & 3 & 34 & -2.4 &
-2.7 & 0.15 & 0.15 \\

2 & $1.0 \times 10^8$ & & 0.058 & $1.0 \times 10^4$ & 1 & 330 & -3.6 &
-3.1 & 0.060 & 0.046 \\

3 & $8.3 \times 10^7$ & & 0.072 & $7.5 \times 10^4$ & 4 & 16 & -2.7 &
-2.9 & 0.052 & 0.077 \\

4 & $8.1 \times 10^7$ & & 0.079 & $1.4 \times 10^5$ & 2 & 38 & -2.6 &
-2.7 & 0.041 & 0.067 \\

5 & $8.1 \times 10^7$ & & 0.058 & $6.8 \times 10^3$ & 4 & 660 & -3.6 &
-2.3 & 0.073 & 0.051 \\

6 & $5.8 \times 10^7$ & & 0.099 & $1.3 \times 10^4$ & 2 & 340 & -3.7 &
-2.6 & 0.047 & 0.082 \\

7 & $3.8 \times 10^7$ & & 0.045 & $2.2 \times 10^3$ & 1 & 440 & -3.8 &
-2.9 & 0.041 & 0.037 \\

8 & $3.0 \times 10^7$ & & 0.13 & $0.0$ & 1 & $\infty$ & -2.1 & -2.8 & 0.21 &
0.82 \\

9 & $3.8 \times 10^7$ & & 0.082 & $1.1 \times 10^4$ & 1 & 110 & -3.5 &
-3.3 & 0.082 & 0.11 \\

\hline
\end{tabular}
\medskip

Column (1): Halo number.  Column (2): Total halo mass.  Column (3):
Best fit $a_c$ in the halo mass accretion history,
eq. (\ref{eqn:mhalo}).  Column (4): Gas fraction.  Column (5): Pop II
stellar mass.  Column (6): Number of Pop III stars hosted in the
progenitors of the halo.  Column (7): Mass-to-light ratio.  Columns
(8,9): Mean metallicity from Pop II SNe and PISNe, respectively.
Columns (10,11): Spin parameters of the gas and DM, respectively.

\end{minipage}
\end{table*}