### 2011-06-23 ApJ/ms.tex
Started to incorporate more intro and methods to the expanded ApJ edition.

\documentclass[apjl]{emulateapj}
%\documentclass[12pt,preprint]{aastex}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{natbib}
%\usepackage[section] {placeins}
\bibliographystyle{apj}
\citestyle{apj}

%%%%%%%% Begin custom definitions %%%%%%%%%%%%%

\newcommand{\fesc}{\ifmmode{f_{\rm esc}}\else{$f_{\rm esc}$}\fi}
\newcommand{\fescs}{\ifmmode{f_{\rm esc}^\star}\else{$f_{\rm esc}^\star$}\fi}
\newcommand{\kms}{\ifmmode{{\rm km~s^{-1}}}\else{km s$^{-1}$}\fi}
\newcommand{\cubecm}{\ifmmode{{\rm cm^{-3}}}\else{cm$^{-3}$}\fi}
\newcommand{\lsim}{\lower0.3em\hbox{$\,\buildrel <\over\sim\,$}}
\newcommand{\gsim}{\lower0.3em\hbox{$\,\buildrel >\over\sim\,$}}
\newcommand{\li}{\noindent$\bullet$\quad}
\newcommand{\flux}{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$}
\newcommand{\emis}{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$ sr$^{-1}$}
\newcommand{\sfr}{M$_\odot$ yr$^{-1}$ Mpc$^{-3}$}
\newcommand{\hst}{{\it HST}}
\newcommand{\enzo}{{\sl Enzo}}
\newcommand{\Ms}{\ifmmode{M_\odot}\else{$M_\odot$}\fi}
\newcommand{\mmin}{M$_{min}$}
\newcommand{\hh}{H$_2$}
\newcommand{\Ol}{$\Omega_\Lambda$}
\newcommand{\Om}{$\Omega_M$}
\newcommand{\Ob}{$\Omega_b$}
\newcommand{\theat}{$t_{\rm{heat}}$}
\newcommand{\tcool}{$t_{\rm{cool}}$}
\newcommand{\rcool}{$r_{\rm{cool}}$}
\newcommand{\tcross}{$t_{\rm{cross}}$}
\newcommand{\tdyn}{$t_{\rm{dyn}}$}
\newcommand{\tkh}{$t_{\rm{KH}}$}
\newcommand{\tH}{$t_{\rm{H}}$}
\newcommand{\tvir}{\ifmmode{T_{\rm{vir}}}\else{$T_{\rm{vir}}$}\fi}
\newcommand{\mvir}{\ifmmode{M_{\rm{vir}}}\else{$M_{\rm{vir}}$}\fi}
\newcommand{\rvir}{\ifmmode{r_{\rm{vir}}}\else{$r_{\rm{vir}}$}\fi}
\newcommand{\rr}{$r_{200}$}
\newcommand{\lya}{Ly$\alpha$}
\newcommand{\jj}{\ifmmode{J_{21}}\else{$J_{21}$}\fi}
\newcommand{\flw}{\ifmmode{F_{LW}}\else{$F_{LW}$}\fi}
\newcommand{\kph}{\ifmmode{k_{\rm ph}}\else{$k_{\rm ph}$}\fi}
\newcommand{\tv}{$\langle T \rangle_{\rm v}$}
\newcommand{\tm}{$\langle T \rangle_{\rm m}$}
\newcommand{\msun}{{\rm\,M_\odot}} 
\newcommand{\lsun}{{\rm\,L_\odot}}
\newcommand{\zsun}{{\rm\,Z_\odot}}
\newcommand{\etal}{et al.\ }
\newcommand\tento[1]{$10^{#1}$}
\newcommand\step[1]{\textit{Step #1}.---}

%%%%%%%% End custom definitions %%%%%%%%%%

\begin{document}

\shorttitle{THE BIRTH OF A GALAXY}
\shortauthors{WISE ET AL.}

\title{The Birth of a Galaxy: Primordial Metal Enrichment and Stellar
  Populations}

\author{John H. Wise\altaffilmark{1,2}, 
  Matthew J. Turk\altaffilmark{3},
  Michael L. Norman\altaffilmark{3},
  Tom Abel\altaffilmark{4}}

\altaffiltext{1}{Department of Astrophysical Sciences, Princeton
  University, Peyton Hall, Ivy Lane, Princeton, NJ 08544}
\altaffiltext{2}{Hubble Fellow}
\altaffiltext{3}{Center for Astrophysics and Space Sciences,
  University of California at San Diego, La Jolla, CA 92093}
\altaffiltext{4}{Kavli Institute for Particle Astrophysics and
  Cosmology, Stanford University, Menlo Park, CA 94025}
\email{jwise@astro.princeton.edu}

\begin{abstract}

  Population III stars first form in dark matter halos with masses
  around $10^6 \Ms$.  By definition, they are metal-free, and their
  protostellar collapse is driven by molecular hydrogen cooling in the
  gas-phase, leading to a massive characteristic mass $\sim 100~\Ms$
  and suppressed fragmentation.  Population II stars with lower
  characteristic masses form when the star-forming gas reaches a
  critical metallicity of $10^{-6} - 10^{-3.5}~Z_\odot$, depending on
  whether dust cooling is important.  We present adaptive mesh
  refinement radiation hydrodynamics simulations that follows the
  transition from Population III to II star formation.  We model
  stellar radiative feedback with adaptive ray tracing.  A top-heavy
  initial mass function for the Population III stars is considered,
  resulting in a plausible distribution of pair-instability supernovae
  and associated metal enrichment.  We find that the gas fraction
  recovers from 5 percent to nearly the cosmic fraction in halos with
  merger histories rich in halos above $10^7 \Ms$.  A single
  pair-instability supernova is sufficient to enrich the host halo to
  a metallicity floor of $10^{-3} Z_\odot$ and to transition to
  Population II star formation.  This provides a natural explanation
  for the observed floor on damped Lyman alpha (DLA) systems
  metallicities reported in the literature, which is of this order.
  We find that stellar metallicities do not necessarily trace stellar
  ages, as mergers of halos with established stellar populations can
  create superpositions of $t-Z$ evolutionary tracks.  A bimodal
  metallicity distribution is created after a starburst occurs when
  the halo can cool efficiently through atomic line cooling.

\end{abstract}

\keywords{cosmology --- methods: numerical --- hydrodynamics ---
  radiative transfer --- star formation}

\section{Motivation}

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \citep{ABN02, Bromm02_P3, OShea07a}.  A fraction of these
stars enrich the surrounding intergalactic medium (IGM) when they go
supernova, which can happen in stars $\lsim 40~\Ms$ in Type II
supernovae (SNe) or in stars roughly between 140 \Ms~and 260 \Ms~in
pair-instability SNe \citep[PISNe;][]{2002ApJ...567..532H}.  The host
halo and the neighboring halos are then enriched with this ejecta.
There exists a critical metallicity that is $\sim 10^{-6} Z_\odot$ if
dust cooling is efficient \citep{Omukai05, Schneider06_Frag, clark08}
and $\sim 10^{-3.5} Z_\odot$ otherwise \citep{Bromm01,
  2009ApJ...691..441S}, where the gas can cool rapidly, lowering its
Jeans mass.  An intermediate characteristic mass of $\sim10~\Ms$ can
be occur if the gas cooling is suppressed to the cosmic microwave
background (CMB) temperature \citep{Larson98, Tumlinson07_IMF,
  2009ApJ...691..441S}.  The resulting Population II star cluster will
thus have a lower characteristic stellar mass than its metal-free
progenitors.  These first stellar clusters may be connected to stars
in the Milky Way halo and nearby dwarf spheroidal (dSph) galaxies,
both with a metallicity floor of [Z/H] = --4 \citep{Beers05,
  Tafelmeyer10, Frebel10_Obs}.

Pop III stars primarily form in dark matter halos with $M \sim 10^6
\Ms$.  Their stellar radiation ionizes and heats an \ion{H}{2} region
with a radius 1--3 kpc.  The overpressurized \ion{H}{2} region drives
a 30 \kms~shock that is 10 times greater than the escape velocity of
the halo, leaving behind a warm ($3 \times 10^4$ K) and diffuse (0.5
\cubecm) medium \citep{Kitayama04, Whalen04, Abel07}.  This aids in
the expansion of the blastwave because it delays the transition to the
Sedov-Taylor and snowplow phases.  In PISNe, approximately half of the
metals stay in the IGM with a metal bubble size of a few kpc
\citep{Wise08_Gal, Greif10}.  The blastwave may induce SF in nearby
halos through the compression of the gas \citep{Ferrara98}, and
timescales for metal mixing into the dense gas are many dynamical
times \citep{Cen08} for shock velocities $\lsim100~\kms$.

The transition from Pop III to Pop II star formation (SF) is solely
dependent on the propagation of metals from the SNe remnants into
future sites of SF.  Their flows are complex because of the
interactions between the SN blastwave, cosmological accretion and halo
mergers, and nearby stellar feedback.  Spurred by the notion of a
critical metallicity, this transition has been extensively studied
with (i) volume-averaged semi-analytic models \citep{Scannapieco03,
  Yoshida04, Furlanetto05_Reion}, (ii) models using hierarchical
merger trees \citep{Tumlinson06, Tumlinson10, Salvadori07, Komiya10},
(iii) post-processing of cosmological simulations with blastwave
models \citep{Karlsson08, Trenti09, Trenti10}, and (iv) direct
numerical simulations with stellar feedback \citep{Tornatore07,
  Ricotti08, Maio10_Pop32, Maio10_Enrich}.

Numerical simulations are useful to detangle and study these
complexities and the transition from Pop III to II stars.  Here, we
present a simulation that includes both types of SF, and their
radiative and mechanical feedback.  The methods used here incorporate
and link together recent results from metal-enriched and metal-free
star formation, the critical metallicity, and pair-instability
supernovae.  This is the first time it has been possible to link the
formation and feedback of the first stars to protogalaxies, resolving
the important scales and including the most important physical
effects.

\section{Method}
\label{sec:setup}

\subsection{Simulation setup}

We use the adaptive mesh refinement (AMR) code
\textsc{enzo~v2.0}\footnote{\texttt{enzo.googlecode.com, changeset
    b86d8ba026d6}} \citep{OShea2004}.  It uses an $N$-body adaptive
particle-mesh solver \citep{Efstathiou85} to follow the dark matter
(DM) dynamics.  It solves the hydrodynamical equation using the
second-order accurate piecewise parabolic method \citep{Woodward84,
  Bryan95}, while a Riemann solver ensures accurate shock capturing
with minimal viscosity.  We use the recently added HLLC Riemann solver
\citep{Toro94_HLLC} for additional stability in strong shocks and
rarefaction waves.  We use the nine-species (\ion{H}{1}, \ion{H}{2},
\ion{He}{1}, \ion{He}{2}, \ion{He}{3}, e$^-$, \hh, \hh$^+$, H$^-$)
non-equilibrium chemistry model in \enzo~\citep{Abel97, Anninos97} and
the \hh~cooling rates from \citet{Glover08_Rates}.  We spatially
distinguish metal enrichment from Pop II and Pop III stars.  We will
follow-up this study with one that considers radiative cooling from
metal-enriched gas, using the method of \citet{2008MNRAS.385.1443S}
that is already implemented in \enzo.

To resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use
a simulation box of 1 Mpc that has a resolution of $256^3$.  This
gives us a DM mass resolution of 1840 \Ms.  We refine the grid on
baryon overdensities of $3 \times 2^{-0.2l}$, where $l$ is the AMR
level, resulting in a super-Lagrangian behavior \citep[also
  see][]{OShea08}.  We also refine on a DM overdensity of $3 \times
2^l$ and always resolve the local Jeans length by at least 4 cells,
avoiding artificial fragmentation during gaseous collapses
\citep{Truelove97}.  If any of these critera are met in a single cell,
it is flagged for further spatial refinement.

We initialize the simulation with \textsc{grafic}
\citep{Bertschinger01} at $z = 130$ and use the cosmological
parameters from the 7-year WMAP $\Lambda$CDM+SZ+LENS best fit
\citep{WMAP7}: $\Omega_M = 0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b
= 0.0449$, $h = 0.71$, $\sigma_8 = 0.81$, and $n = 0.963$ with the
variables having their usual definitions.  We use a maximum refinement
level of $l = 12$, resulting in a maximal comoving resolution of 1 pc.
We stop the simulation at $z=7$ to prevent any large-scale modes with
$r \sim L_{\rm box}/2$ from entering the non-linear growth regime.  At
the final redshift, the baseline simulation has $1.4 \times 10^8$
computational cells and required 250,000 CPU hours, using 512 cores.

\subsection{Star formation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
%\epsscale{1.15}
\plotone{f1}
%\epsscale{1}
\caption{\label{fig:massfn} \textit{Top}: Halo mass function of the
  simulation compared with the analytical fit (solid) of
  \citet{Warren06}.  \textit{Bottom}: Total Pop II luminosities of
  halos with constant mass-to-light ratios overplotted.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}
%\epsscale{1.15}
\plotone{f2}
%\epsscale{1}
\caption{\label{fig:projections} Density-weighted projections of gas
  density (top), temperature (middle), and metallicity (bottom) at
  $z=7$.  The left column shows the entire simulation volume, where
  the center and right columns focus on the intense and quiet halos,
  which are marked by left and right arrows in the upper-left panel.
  The metallicity projections are a composite picture of metals
  originating from Pop III (red) and Pop II (blue) stars with magenta
  indicating a mixture of the two.}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We distinguish Pop II and Pop III SF by the total metallicity of the
densest cell in the molecular cloud.  Pop II stars are formed if [Z/H]
$> -4$, and Pop III stars are formed otherwise.  We do not consider
Pop III.2 stars and intermediate mass stars from CMB-limited cooling.

Simulations have shown that the characteristic mass of Pop III stars
$M_{\rm char} \sim 100~\Ms$.  They form in molecular clouds that
coexist with the dark matter halo center with limited fragmentation
occurring during their collapse; however \citet{2009Sci...325..601T}
and \citet{Stacy10_Binary} have recently shown that Pop III binaries
may form in a fraction of such halos.

% Furthermore, detailed one-dimensional calculations have shown that
% their IMF should follow a Kroupa-like IMF that has a Salpeter slope
% at the high-mass end and an exponential cut-off below $M_{\rm char}$
% \cite{refs}.

For Pop III stars, we use the same SF model as \citet{Wise08_Gal} where
each star particle represents a single star, forming at an overdensity
of $5 \times 10^5$.  Instead of using a fixed stellar mass, we randomly
sampled from an IMF with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm
      char}}{M}\right)^{1.6}\right] dM
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we record the number
of times the random number generator (Mersenne twister;
\citet{MTwister}) has been called for use when restarting the
simulations.

%%% FOR THE COMPARISON BETWEEN RUNS WITH AND WITHOUT METAL COOLING %%%
%
% Although the Pop III star formation history (SFH) is identical for a
% given $N$ stars, the Pop III SFH for a given halo can be different
% between the two simulations.  This happens because the order in which
% halos form Pop III stars can be altered from enhanced star formation
% and feedback from nearby Pop II star clusters, as we will demonstrate
% in this Letter.

We treat Pop II star formation with the same prescription as
\citet{Wise09}, which is similar to the Pop III prescription but
without the \hh~fraction floor that is removed because the
metal-enriched gas can efficiently cool even in the presence of a
strong UV radiation field \citep[e.g.][]{Safranek10}.  This method is
effectively a modified version of the \citet{Cen92} method but
accounts for the fact that the molecular clouds are resolved.  In the
simulation, we define the molecular cloud as a sphere, centered on the
most dense cell, with a dynamical time $t_{\rm dyn} = 3$ Myr
(corresponding to an average density $\bar{\rho}_{\rm MC} \simeq
1000\mu \cubecm$) and a radius $R_{\rm MC}$, where $\mu$ is the mean
molecular weight.  In this sphere, a fraction $c_\star = 0.07 f_{\rm
  cold}$ of the cold gas ($T < 10^3$ K) is converted into a star
particle with mass $m_\star = c_\star (4\pi/3) \bar{\rho}_{\rm MC}
R_{\rm MC}^3$.  We then replace the sphere with a uniform density
$\rho_{\rm MC} = (1 - c_\star) / (Gt_{\rm dyn}^2)$ and temperature $T
= 10^4$ K, which approximates the initial stages of an \ion{H}{2}
region.  Similar to the Pop III treatment, we merge any newly created
particles within a radius $R_{\rm merge}$.  We set the minimum mass of
a star particle to $m_{\rm \star, min} = 1000 \Ms$.  If the initial
mass does not exceed $m_{\rm \star, min}$, the star particle does not
provide any feedback and continues to accrete until it reaches $m_{\rm
  \star, min}$.  The star particles live for 20 Myr, the maximum
lifetime of an OB star.  These stars generate the majority of the
ionizing radiation and SNe feedback in stellar clusters, thus we
ignore any feedback from lower mass stars.

\subsection{Stellar feedback}

We use mass-dependent luminosities and lifetimes of the Pop III stars
from \citet{Schaerer02}.  The radiation field is evolved with adaptive
ray tracing \citep{Abel02_RT, Wise10} and is coupled self-consistently
to the hydrodynamics.  We model the \hh~dissociating radiation with an
optically-thin, inverse square profile, centered on all star
particles.  These stars die as pair-instability SNe (PISNe) if they
are in the mass range 140--260 \Ms~\citep{Heger03}.  We use the
explosion energy from \citet{Heger02}, where we fit the following
function to their models, $E_{\rm PISN} = 10^{51} \times [5.0 + 1.304
  (M_{\rm He} - 64)]$, where $M_{\rm He} = (13/24) \times (M_\star -
20) \Ms$ is the helium core (and equivalently the ejecta) mass and
$M_\star$ is the stellar mass.

The Pop II star particles emit 6000 hydrogen ionizing photons per
baryon over their lifetime, and we do not consider singly- and
doubly-ionizing helium photons.  We note that low-metallicity stars
generate up to a factor of four more ionizing photons than a solar
metallicity population \citep{Schaerer03} and might be underestimating
the radiative feedback.  Nonetheless this study provides an excellent
first insight in the transition to Pop II SF, as the metal enrichment
is the key ingredient here.  For SN feedback, these stars generate
$6.8 \times 10^{48}$ erg s$^{-1}$ $\Ms^{-1}$ after living for 4 Myr,
which is injected into spheres with a radius of 10 pc.  If the
resolution of the grid is less than 10/3 pc, we deposit the energy
into a $3^3$ cube surrounding the star particle.

\begin{figure*}
%\epsscale{1.15}
\plotone{f3}
%\epsscale{1}
\caption{\label{fig:evo} (a) Evolution of the total halo mass (top),
  stellar mass (middle), and gas fraction (bottom) of the quiet
  (dashed) and intense (solid) halos.  (b) Mass-weighted stellar
  metallicities and gas metallicities enriched by Pop II and Pop III
  SNe of the intense (top) and quiet (bottom) halos.}
\end{figure*}

\section{Results}
\label{sec:results}

Here we present the gaseous and stellar evolution of two selected
halos in the simulation: one that has an early mass buildup but no
major mergers after $z=12$, and one that experiences a series of major
mergers between $z=10$ and $z=7$.  We name the halos ``quiet'' and
``intense'', respectively.  The entire simulation contains 38 galaxies
with 3640 Pop II stellar clusters and captures the formation of 333
Pop III stars.  The halo mass function and galaxy luminosities are
plotted in Figure \ref{fig:massfn}.

% We start with the mass accretion history and enrichment of the
% halos.  We then compare the nature of the stellar populations in
% these halos.

We illustrate the state of the simulation at $z=7$ in Figure
\ref{fig:projections} with density weighted projections of gas
density, temperature, and metallicity, showing the entire box and
focusing on the two halos of interest.  Radiative and mechanical
feedback create a multi-phase medium inside these halos, which are
embedded in a warm and ionized IGM.

\subsection{Mass evolution}
\label{sec:halo}

%\li Describe the evolution of the baryon fraction and metallicity from
%PISN and Type II SNe metals in the two halos.

Figure \ref{fig:evo}a shows the total, metal-enriched stellar, and gas
mass history of the most massive progenitors of both halos.  The quiet
halo undergoes a series of major mergers at $z > 12$, growing by a
factor of 30 to $2.5 \times 10^7 \Ms$ within 150 Myr.  Afterwards it
only grows by a factor of 3 by $z=7$ mainly through smooth accretion
from the filaments and IGM.  It is the most massive halo in the
simulation between redshifts 13 and 10.  At the same time, the intense
halo has a mass $M = 3 \times 10^6 \Ms$, but it is contained in a
biased region on a comoving scale of 50 kpc with $\sim25$ halos with
$M \sim 10^6 \Ms$.  After $z=10$, these halos hierarchically merge to
form a $10^9 \Ms$ halo at $z=7$ with two major mergers at redshifts 10
and 7.9, seen in the rapid increases in total mass in Figure
\ref{fig:evo}a.  The merger history of the two halos are not atypical
as dark matter halos can experience both quiescent and vigorous mass
accretion rates.

Both halos start forming Pop II stars when $M = 10^7 \Ms$.  This is
consistent with the filtering mass $M_f$ of high-redshift halos when
it accretes mainly from a pre-heated IGM \citep{gnedin98, gnedin00,
  Wise08_Gal}.  Afterwards these halos can cool efficiently through
\hh~cooling, sustaining constant and sometimes bursting SF.  The
latter characteristics are equated with the definition of a galaxy.
The quiet halo forms $10^5 \Ms$ of stars by $z=9$.  This initial
starburst photo-evaporates the majority of its molecular clouds, in
addition to heating and ionizing the surrounding IGM out to a radius
of 10--15 kpc at $z=9$.  These respectively reduce the in-situ and
external cold gas supply that could feed future SF.

The gas fractions of both halos decrease from 0.15 to 0.08 by outflows
driven by ionization fronts and blastwaves in their initial
starbursts.  The quiet halo does not have a major merger with any halo
with $M > M_f$, leading to a small final gas fraction.  These low-mass
halos are photo-evaporated, hosting diffuse warm gas reservoirs
instead of cold dense cores.  After $z=10$, the halo mainly accretes
warm diffuse gas from the filaments and IGM.  In contrast, the intense
halo grows from major mergers of halos with $M > M_f$.  The progenitor
halos involved in the major mergers are able to host molecular clouds
and have higher gas fractions.  Between $z=10$ and $z=8$, the gas
fraction increases from 0.07 to 0.12 until it jumps to 0.14 when a
gas-rich major merger occurs.  The stellar mass accordingly increases
with the ample supply of gas during this period.

\subsection{Metallicity evolution}

The evolution of the stellar and gas metallicity of both halos are
illustrated in Figure \ref{fig:evo}b.  PISNe from Pop III stars enrich
the nearby IGM out to a radius of 10 kpc and provides a metallicity
floor of $[Z_3/H] \sim -3$.  Ejecta from Pop II SNe initially enrich
the ISM of both halos to an average $[\bar{Z}_2/H]$ between --3.5 and
--3.

In the quiet halo, an equilibrium of $[\bar{Z}_2/H] \sim -2.5$ is
established between metal-rich outflows and metal-poor inflows.
Galactic outflows are directed in the polar directions of the gas
disk, keeping the adjacent filaments metal-poor.  These features and a
well-mixed ISM \citep[cf.][]{Wise08_Gal, Greif10} are apparent in the
metallicity projections in Figure \ref{fig:projections}.  The average
stellar metallicity is within 0.5 dex of $[\bar{Z}_2/H]$.

In the intense halo, the first few Pop II star clusters have $[Z/H]$
between --1 and --2 and dominate the average stellar metallicity at $z
> 8$.  Afterwards the metallicity increases by a factor of 30 to
$[\bar{Z}_2/H] = -1.5$ through self-enrichment from a starburst.
Because this halo is located in a large-scale overdensity, most of the
ejecta falls back into the halo after reaching distances up to 20
comoving kpc, keeping the halo gas metallicity high because the
inflows are relatively metal-rich themselves.  After the $z \sim 8$
starburst, the average stellar metallicity follows the average gas
metallicity within 0.1 dex.

\begin{figure*}
%\epsscale{1.1}
\plotone{f4}
%\epsscale{1}
\caption{\label{fig:pops} The scatter plots show the SF history of the
  quiet (left) and intense (right) halos as a function of metallicity
  at $z=7$.  Each circle represents a star cluster, whose area is
  proportional to its mass.  The open circles in the upper right
  represent $10^3$, $10^4$, and $10^5$ \Ms~star clusters.  The dashed
  lines in the right panel guide the eye to two stellar populations
  that were formed in two satellite halos, merging at $z=7.5$.  The
  upper histogram shows the SFR.  The right histogram depicts the
  stellar metallicity distribution.}
\end{figure*}

\subsection{Star formation history}
\label{sec:pop}

The most massive progenitor of the quiet halo interestingly never
hosts a Pop III star.  Instead a nearby halo forms a Pop III star,
which (randomly) produces a PISNe at $z=16$.  The blastwave overruns
the most massive progenitor, and the dense core survives this event
and is enriched by this PISN, triggering the transition to Pop II SF.
Other progenitors host three Pop III stars, forming at $z = 15.4,
14.2, 13.8$, with the latter producing a PISN.  Metal enrichment from
these two PISNe and Pop II SF quench Pop III SF in this halo.  The
progenitors of the intense halo host a total of 56 Pop III stars with
21 producing PISNe.  The first forms in a $6 \times 10^5 \Ms$ halo at
$z=19$.  Pop III stars form on a regular interval in the halo's
progenitors until $z=9$ when most of these halos enter the metal-rich
bubble surrounding the intense halo.

%% \li Present results on the star formation history and the physical
%% reasons for the features seen in the SFH and metallicities.
%%
%% \li Present the stellar metallicity distribution at the final time

Figure \ref{fig:pops} shows the SF history (SFH), metallicity
distribution, and SF rates of both halos.  A nearby PISNe provides a
metallicity floor of [Z/H] = --2.8 in the quiet halo at which
metallicity the first Pop II stars form.  The stellar metallicity
evolution exhibits what is expected from an isolated system with the
stellar feedback steadily enriching the ISM, resulting in a
correlation between stellar age and metallicity.  After $z=10$ the
metallicities plateau at [Z/H] = --2.1 for reasons previously
discussed.  The SFR peaks at $z=10$ and decreases as the cold gas
reservoir is depleted.  Around $z=7.5$, a 25:1 minor merger occurs,
and the gas inside the satellite halo is compressed, triggering
metal-poor, [Z/H] = --3.2, SF during its nearest approach.  This halo
remains metal-poor because most of metal enrichment in the quiet halo
occurs in bi-polar flows perpendicular to the galaxy disk and
filament.  Stars with [Z/H] $<$ --3 compose 1.6 percent of the total
stellar mass.

In contrast with the quiet halo, the intense halo undergoes a few
mergers of halos with an established stellar population.  This creates
a superposition of age-metallicity tracks in the SFH, seen in the
complexity of Figure \ref{fig:pops}.  The first two Pop II stellar
clusters have an unexpectedly high metallicity [Z/H] $\sim$ --1, which
occurs when a PISN blastwave triggers SF in two neighboring halos.
Most of the early SF have [Z/H] = --2.5.  At $z=9$, the halo's virial
temperature reaches \tento{4} K.  This combined with a 10:1 merger
creates a starburst that quickly enriches the halo to [Z/H] = --1.5 by
$z=8$.  The halo continues to enrich itself afterwards.  The spikes in
the scatter plot correspond to SN triggered SF in nearby molecular
clouds that are enriched up to a factor of 10 with respect to the ISM.
However their mass fraction are small compared to the total stellar
mass.  The starburst at $z=9$ creates a bimodal metallicity
distribution with peaks at [Z/H] = --2.4 and --1.2 with the metal-rich
component mainly being created after the starburst.  Two systems with
sizable stellar components merge into the halo at $z \sim 8$, and
their stellar populations are still discernible in the metallicity-age
plot.  Stars with [Z/H] $<$ --3 compose 1.8 percent of the total
stellar mass.

\section{Discussion and Conclusions}

In this Letter, we focus on the birth of two galaxies prior to
reionization with a cosmological AMR radiation hydrodynamics
simulation.  Supernovae from Pop III stars provide the necessary heavy
elements for the transition to a Population II stellar population,
which we have directly simulated.  These two galaxies have a 10--15\%
probability in surviving as present-day ``fossil'' galaxies
\citep{Gnedin06}, otherwise they will be incorporated into galactic
stellar halos.  A $z=8.55$ galaxy was recently spectroscopically
confirmed that is contained in an ionized bubble with radius 0.1--0.5
Mpc with an uncharacterized population of galaxies contributing to the
local ionizing radiation field \citep{Lehnert10_z8.6}.  The galaxies
simulated here might represent this population and should be detected
with the upcoming \textit{James Webb Space Telescope}.

We find that one PISN is sufficient to enrich the star-forming halo
and surrounding $\sim 5$ kpc to a metallicity of 10$^{-3} Z_\odot$,
given $M_{\rm char} = 100~\Ms$.  DLA systems have a metallicity floor
on the same order \citep{Wolfe05_Review}, and metal enrichment from Pop
III SNe provides a possible explanation.

If the first stars have a lower characteristic mass that favor
hypernovae \citep{Tumlinson07_IMF}, then this metallicity floor should
be lowered by a factor of $\sim 10$ because (1) the metal ejecta is
lowered by a factor of $\sim 50$ and (2) the mixing mass is
approximately decreased by a factor of $(E_{\rm hyp}/E_{\rm
  PISN})^{3/5}$ in the Sedov-Taylor solution, where $(E_{\rm
  hyp}/E_{\rm PISN}) \sim 0.1$ is the ratio of explosion energies of a
hypernova and PISN.  In the case where this metallicity floor is less
than the critical metallicity, then the next instance of SF will
further enrich the ISM, solidifying the transition to Pop II SF.  We
conclude that it only takes one, at most two, SNe from Pop III stars
in the halo progenitors to complete the transition to Population II
\citep{Frebel10}.  The question of whether the critical metallicity is
\tento{-6} or \tento{-3.5} $Z_\odot$ is most applicable to nearby
halos where the heavy elements mix slowly into dense cores as the
blastwave overtakes it.  Less than two percent of stars have [Z/H] $<
-3$ in both systems, consistent with observations of metal-poor stars
in the halo and dSphs \citep{Beers05, Battaglia10}.

% We also find that the merger history plays an important role in
% supplying gas into halos after its first epoch of SF.  Mergers of
% halos below the filtering mass are inefficient in providing gas
% whereas the opposite is true for merging halos above the filtering
% mass.  Halos do not necessarily need $\tvir \ge 10^4$ K to form
% significant stellar populations; however SFRs dramatically increase,
% and thus metal enrichment, when this threshold is reached.  Halos
% with mostly gas-poor mergers or a quiet merger history result in a
% monotonic increase in metallicity with time.  SFHs become more
% complex with multiple metallicity-age tracks if the halo experiences
% mergers with other halos that have an established stellar
% population.  Furthermore SF that is triggered by blastwaves
% interacting with molecular clouds can have metallicities up to a
% factor of 10 higher than the main starburst.  Starbursts at $\tvir =
% 10^4$ K can enrich the host halo enough to create a bimodal
% metallicity distribution, where the metal-poor component is created
% before the burst.  Note that mergers of stellar populations can also
% create a similar bimodal distribution, which are observed in dSphs
% \cite{Battaglia10}.

We have shown that it is possible to simulate the formation of a
high-redshift dwarf galaxy and its entire SFH with radiative and
mechanical feedback.  These results provide invaluable insight on the
first galaxies and the role of metal-free stars in the early universe.
There exists a wealth of information in this simulation, and we plan
to follow up this preliminary report with more detailed analysis of
the metal enrichment of the IGM, global SF rates, and observational
connections with high-redshift galaxies and Local Group dwarf galaxies
in the near future.

\acknowledgments

J.~H.~W. is supported by NASA through Hubble Fellowship grant
\#120-6370 awarded by the Space Telescope Science Institute, which is
operated by the Association of Universities for Research in Astronomy,
Inc., for NASA, under contract NAS 5-26555.  M.~J.~T. and
M.~L.~N. acknowledge partial support from NASA ATFP grant NNX08AH26G.
Computational resources were provided by NASA/NCCS award SMD-09-1439.
J.~H.~W. thanks Renyue Cen, Amina Helmi, Marco Spaans, and Eline
Tolstoy for enlightening discussions.  The majority of the analysis
and plots were done with \texttt{yt} \citep{yt_full_paper}.

%\clearpage
%\input biblio.tex
\bibliography{ms}
\end{document}