### 2011-08-04 ApJ/table1.tex

\begin{deluxetable*}{ccccccc}
%\tablecolumns{7}
\tabletypesize{}
\tablewidth{0pc}
\tablecaption{Enrichment Fractions\label{tab:enrich}}

\tablehead{\colhead{Source} & \colhead{Weighting} & 
  \multicolumn{5}{c}{Metallicity Cut (in units of [Z/H])} \\
%%%%%%%%%%
  \cline{3-7} \\
%%%%%%%%%%
  \colhead{} & \colhead{} & 
  \colhead{-6} & \colhead{-5} & \colhead{-4} & \colhead{-3} & \colhead{-2}
}
\startdata
Pop III & Volume & 0.0384 & 0.0331 & 0.0268 & 0.0188 & 0.00836 \\
Pop III & Mass & 0.144 & 0.125 & 0.102 & 0.0650 & 0.0124 \\
Pop II & Volume & 0.00560 & 0.00397 & 0.00236 & 0.00102 & 0.000266 \\
Pop II & Mass & 0.0525 & 0.0438 & 0.0344 & 0.0251 & 0.0174
\enddata
\end{deluxetable*}