### 2011-08-05 ApJ/table1.tex
Reworked enrichment fraction table.

\begin{deluxetable}{ccccc}
%\tablecolumns{7}
\tabletypesize{}
\tablewidth{\columnwidth}
\tablecaption{Enrichment Fractions\label{tab:enrich}}

\tablehead{\colhead{[Z/H]} & 
  \colhead{$f_v$(Pop III)} & \colhead{$f_m$(Pop III)} & 
  \colhead{$f_v$(Pop II)} & \colhead{$f_m$(Pop II)}
}
\startdata
-6 & 3.84(-2) & 1.44(-1) & 5.60(-2) & 5.25(-2) \\
-5 & 3.31(-2) & 1.25(-1) & 3.97(-2) & 4.38(-2) \\
-4 & 2.68(-2) & 1.02(-1) & 2.36(-2) & 3.44(-2) \\
-3 & 1.88(-2) & 6.50(-2) & 1.02(-2) & 2.51(-2) \\
-2 & 8.36(-3) & 1.24(-2) & 2.66(-3) & 1.74(-2)
\enddata

\tablecomments{$f_v$ and $f_m$ are volume filling and mass-weighted
  fractions, respectively, above several different metallicity cuts.
  Values are listed in scientific notation with the exponents in
  parentheses.}

\end{deluxetable}