### 2011-09-08 ApJ/table1.tex

\begin{deluxetable}{ccccc}
\tablecolumns{5}
\tabletypesize{}
\tablewidth{\columnwidth}
\tablecaption{Enrichment Fractions at Redshift 7\label{tab:enrich}}

\tablehead{\colhead{[Z/H]} & 
  \colhead{$f_v$(Pop III)} & \colhead{$f_m$(Pop III)} & 
  \colhead{$f_v$(Pop II)} & \colhead{$f_m$(Pop II)}
}
\startdata
-6 & 3.81(-2) & 1.43(-1) & 5.52(-3) & 5.22(-2) \\
-5 & 3.28(-2) & 1.24(-1) & 3.92(-3) & 4.36(-2) \\
-4 & 2.65(-2) & 1.01(-1) & 2.33(-3) & 3.43(-2) \\
-3 & 1.86(-2) & 6.45(-2) & 1.00(-3) & 2.50(-2) \\
-2 & 8.29(-3) & 1.23(-2) & 2.61(-4) & 1.73(-2)
\enddata

\tablecomments{$f_v$ and $f_m$ are volume filling and mass-weighted
  fractions, respectively, above several different metallicity cuts.
  Values are listed in scientific notation with the exponents in
  parentheses.}

\end{deluxetable}